---
title: "Prediction Assignment Writeup"
author: "Juan Tenopala"
date: "19 de diciembre de 2017"
output:
  html_document:
    keep_md: true
---

#Prediction Assignment Writeup

## Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Objective
To use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants, to predict the manner in which they did the exercise.

## Data
The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.
The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

## Introduction
One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.
### The model
The outcome variable is classe, a factor variable with 5 levels. For this data set, participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in 5 different fashions: Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes.
Prediction evaluations will be based on maximizing the accuracy and minimizing the out-of-sample error. All other available variables after cleaning will be used for prediction.
In this case I will use a Random Forest algorithm because the features of the algorithm are important for classification and prediction.

## Code and Results
### Packages, Libraries and seed
```{r}
library(caret)
library(randomForest)
set.seed(1234)
```

### Load the data sets
```{r}
trainingset <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!", ""))
testingset <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!", ""))
dim(trainingset)
dim(testingset)
```

### Preprocessing
Delete columns with all "NA" and the variables that are irrelevant to the analysis.
```{r}
trainingset<-trainingset[,colSums(is.na(trainingset)) == 0]
testingset <-testingset[,colSums(is.na(testingset)) == 0]
trainingset   <-trainingset[,-c(1:7)]
testingset <-testingset[,-c(1:7)]
```

### Exploratory data analysis
```{r}
subsamples <- createDataPartition(y=trainingset$classe, p=0.7, list=FALSE)
subTraining <- trainingset[subsamples, ] 
subTesting <- trainingset[-subsamples, ]
dim(subTraining)
dim(subTesting)
head(subTraining)
head(subTesting)
```

```{r}
plot(subTraining$classe, col="blue", main="5 levels in variable classe of the subTraining data set", xlab="classe levels", ylab="Frequency")
```
From the graph above, we can see that each level frequency is within the same order of magnitude of each other. Level A is the most frequent with more than 4000 occurrences while level D is the least frequent with about 2500 occurrence

### Prediction model
```{r}
model <- randomForest(classe ~. , data=subTraining, method="class")
subprediction <- predict(model, subTesting, type="class")
prediction <- predict(model, testingset, type="class")
prediction
```

